Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
knitr::opts_chunk$set(echo = TRUE)
fb_data <- read.csv("~/R_projects/class_data_analytics/data/pseudo_facebook.tsv", sep = "\t")

```




```{r}
head(fb_data)
```


***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)

qplot(data = fb_data, x=age)
quantile(fb_data$age)

qplot(data = fb_data, x=dob_day) +
   scale_x_continuous(breaks=1:31)

```

***

#### What are some things that you notice about this histogram?
Response: dob on the 1th of a month is very frequent, also the 15th. the 31th seems under-representeted. interpretation: not every month has 31 days, so the under-represention of 31th is realistic. the 1th of a month is not very intuitiv - maybe false input for day of birth? same with 15th

***

### Moira's Investigation
Notes: Moira investigation is about quantification of audience size. her findings is that most users underestimates their size of audience.  

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes: 

```{r Faceting}
qplot(data = fb_data, x = dob_day) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let???s take another look at our plot. What stands out to you here?
Response: the 01.01. is overrepresentated. it seems that my assumpation about fake dob input is more realistic, because its the default input of the dropdown menue

***

### Be Skeptical - Outliers and Anomalies
Notes: 

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira???s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data=fb_data, x= friend_count)
```

#### what is the plot looks like?
Response: the variable distribution is squished into the lower section this distrubtion called "long tail" data

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

qplot(data=fb_data, x= friend_count, binwidth = 40) +
  scale_x_continuous(limits=c(1,1000))

#qplot(data=fb_data, x= friend_count, xlim=c(0,1000), binwidth = 40)
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = fb_data, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

qplot(x = friend_count, data = na.omit(fb_data), binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

table(fb_data$gender)
by(fb_data$friend_count, fb_data$gender, mean)
by(fb_data$friend_count, fb_data$gender, summary)

```

#### Who on average has more friends: men or women?
Response:
mean female: 241.9699
mean male:   165.0355

#### What's the difference between the median friend count for women and men?
Response:
mean female: 96 
mean male:   74 

#### Why would the median be a better measure than the mean?
Response: its robust against extreme values (outlier). median < mean is because of the long tail distribution

***

### Tenure
Notes:

```{r Tenure}
qplot(data = fb_data, x = tenure, binwidth = 30, color = I('black'), fill = I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(data = fb_data, x = tenure/365, binwidth = 0.25, color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous(limits = c(0,7),  breaks = seq(1,7,1))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(data = fb_data, x = tenure/365, binwidth = 0.25, color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous(limits = c(0,7),  breaks = seq(1,7,1))+
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

***

### User Ages
Notes:

```{r User Ages}
qplot(data = fb_data, x = age, binwidth = 2, color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous(limits = c(min(fb_data$age),100), breaks= seq(10, 100, 5))+
  xlab('Age of user') +
  ylab('Number of users in sample')


qplot(data = na.omit(fb_data), x = age, binwidth = 2, color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous(limits = c(min(fb_data$age),100), breaks= seq(10, 100, 5))+
  xlab('Age of user') +
  ylab('Number of users in sample') +
  facet_wrap(~gender)
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:


### Scaling x axis with transforming x

***
```{r}
#install.packages("gridExtra")
library(gridExtra)

# define individual plots
p1 = qplot(data = fb_data, x=friend_count) +
  xlab('number of friends') +
  ylab('occurrance - normal scale')
p2 = qplot(data = fb_data, x=sqrt(friend_count)) +
  xlab('number of friends') +
  ylab('occurrance - sqrt scale')
p3 = qplot(data = fb_data, x=log2(friend_count+1)) + 
  xlab('number of friends') +
  ylab('occurrance - log2 scale')
p4 = qplot(data = fb_data, x=log10(friend_count+1))+
  xlab('number of friends') +
  ylab('occurrance - log10 scale')

# arrange plots in grid
grid.arrange(p1, p2, p3, p4 , ncol=2)
```

### Scaling x axis without transforming x

```{r Add a Scaling Layer}
#install.packages("gridExtra")
library(gridExtra)

# define individual plots
p5 = qplot(data = fb_data, x=friend_count) +
  xlab('number of friends') +
  ylab('occurrance - normal scale')
p6 = qplot(data = fb_data, x=friend_count) +
  scale_x_continuous(trans='sqrt') +
  xlab('number of friends') +
  ylab('occurrance - sqrt scale')
p7 = qplot(data = fb_data, x=friend_count+1) + 
  scale_x_continuous(trans='log2') +
  xlab('number of friends') +
  ylab('occurrance - log2 scale')
p8 = qplot(data = fb_data, x=friend_count+1) +
  scale_x_continuous(trans='log10')+
  xlab('number of friends') +
  ylab('occurrance - log10 scale')

# arrange plots in grid
grid.arrange(p1, p2, p3, p4 , ncol=2)
```

***
### difference:

```{r}

# arrange plots in grid
grid.arrange(p1, p5, p2, p6, p3, p7, p4, p8, ncol=2)
```


### Frequency Polygons

```{r Frequency Polygons}
qplot(data = na.omit(fb_data), x = friend_count, binwidth = 10, color = gender, geom = "freqpoly")+
  scale_x_continuous(limits = c(min(fb_data$friend_count),max(fb_data$friend_count)))+
  xlab('friend_count of user') +
  ylab('Number of users in sample')

qplot(data = na.omit(fb_data), x = friend_count, y=..count../sum(..count..), binwidth = 10, color = gender, geom = "freqpoly")+
  scale_x_continuous(limits = c(min(fb_data$friend_count),max(fb_data$friend_count)))+
  xlab('friend_count of user') +
  ylab('Number of users in sample (relative)')

## grah for friend_count < 500
qplot(data = na.omit(fb_data), x = friend_count, y=..count../sum(..count..), binwidth = 10, color = gender, geom = "freqpoly")+
  scale_x_continuous(limits = c(min(fb_data$friend_count),500))+
  xlab('friend_count of user') +
  ylab('Number of users in sample (relative)')


qplot(data = na.omit(fb_data), x = friend_count, y=..count../sum(..count..), binwidth = 100, color = gender, geom = "freqpoly")+
  scale_x_continuous(limits = c(500 ,1500))+
  xlab('friend_count of user') +
  ylab('Number of users in sample (relative)')

qplot(data = na.omit(fb_data), x = friend_count, y=..count../sum(..count..), binwidth = 100, color = gender, geom = "freqpoly")+
  scale_x_continuous(limits = c(1500 ,3000))+
  xlab('friend_count of user') +
  ylab('Number of users in sample (relative)')

qplot(data = na.omit(fb_data), x = sqrt(friend_count), binwidth = 10, color = gender, geom = "freqpoly")+
  xlab('friend_count of user') +
  ylab('Number of users in sample')

qplot(data = na.omit(fb_data), x = sqrt(friend_count), binwidth = 10, color = gender, geom = "freqpoly")+
  scale_x_continuous(limits = c(30 ,70))+
  xlab('friend_count of user') +
  ylab('Number of users in sample')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(data = na.omit(fb_data), x = www_likes, binwidth = 10, color = gender, geom = "freqpoly")+
  #scale_x_continuous(limits = c(min(fb_data$friend_count),max(fb_data$friend_count)))+
  xlab('www_likes of user') +
  ylab('Number of users in sample')

qplot(data = na.omit(fb_data), x = log10(www_likes+1), binwidth = 0.05, color = gender, geom = "freqpoly")+
  #scale_x_continuous(limits = c(min(fb_data$friend_count),max(fb_data$friend_count)))+
  xlab('www_likes of user') +
  ylab('Number of users in sample')

qplot(data = subset(fb_data, !is.na(gender)), x = www_likes, color = gender, geom = "freqpoly")+
  scale_x_continuous(trans="log10")+
  xlab('www_likes of user') +
  ylab('Number of users in sample')

```

### whats the www_likes count for male? 


```{r}
by(fb_data$www_likes, fb_data$gender, sum)
by(fb_data$www_likes, fb_data$gender, summary)
```

***

### Box Plots
Notes:

```{r Box Plots}
qplot(data=subset(fb_data, !is.na(gender)), x= gender, y=friend_count, geom="boxplot")


```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(data=subset(fb_data, !is.na(gender)), x= gender, y=friend_count, geom="boxplot")+
  scale_y_continuous(limits = c(0,1000))

qplot(data=subset(fb_data, !is.na(gender)), x= gender, y=friend_count, geom="boxplot")+
  coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(data=subset(fb_data, !is.na(gender)), x= gender, y=friendships_initiated, geom="boxplot")
  #coord_cartesian(ylim = c(0,1000))

qplot(data=subset(fb_data, !is.na(gender)), x= gender, y=friendships_initiated, geom="boxplot") +
  coord_cartesian(ylim = c(0,300))

by(fb_data$friendships_initiated, fb_data$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response: female
#### Write about some ways that you can verify your answer.
Response: group column friendship_initiated by gender and take the average
```{r Friend Requests by Gender}
by(fb_data$friendships_initiated, fb_data$gender, mean)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
cond = fb_data$mobile_likes > 0
fb_data$mobile_check_in <- ifelse(cond, 1, 0)
## factorize to define variable as categorical
fb_data$mobile_check_in <- factor(fb_data$mobile_check_in)

summary(fb_data$mobile_check_in) / length(fb_data$mobile_check_in)

```

Response:

***

### Analyzing One Variable
Reflection:
- scale a variable could help to see pattern. therefore we use sqrt or a type of log function
- to explore pattern we use some kind of plots. we have learned to use freqpoly, boxplots and histograms to analyse the distribution of a variable
- sometimes variables distribution are "long tail". with scaling and zooming into plots, more insights could be generated. it could help to transform a variable with lots of outliers in binary or multiple classes  
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!