---
title: "lesson4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load and inspect data  


```{r diamonds}

library(ggplot2) #must load the ggplot package first
data(diamonds) #loads the diamonds data set since it comes with the ggplot package

summary(diamonds)
str(diamonds)
summary(diamonds$color)
?diamonds


```

## Quiz1

How many observations? 
A: 53940

How many vars? 
A: 10

How many ordered factors? 
A: 3

Which is the "best" factor of diamond color? 
A: J


## Quiz2

Plot price distribution. Is it long tail data? 
A: Yes
```{r plot hist of price}

qplot(data = diamonds, x=price, bins=50) 
```

## Quiz3

Describe the distribution: center, mean, std,...


```{r}
summary(diamonds$price)

```

## Quiz4

How many diamonds costs...

less than 500?
less than 250?
more than 15000?
A: 

```{r}
price_under_250 = subset(diamonds, price<250) 
price_under_500 = subset(diamonds, price<500)
price_equalover_15000 = subset(diamonds, price>=15000)

dim(price_under_250)
dim(price_under_500)
dim(price_equalover_15000)

```

## Quiz5

Look at the largest peak of price
There are no diamonds that cost $1500.

For diamonds that cost less than 2,000, the most common price of a diamond is around 700 with the mode being $605

```{r}
qplot(data = diamonds, x = price, binwidth = 30, color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous(limits = c(min(diamonds$price),2000), breaks= seq(0, max(diamonds$price), 1000))+
  xlab('Age of user') +
  ylab('Number of diamonds in sample')
qplot(data = diamonds, x = sqrt(price), binwidth = 1, color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous() +
  xlab('sqrt price of diamonds') +
  ylab('Number of diamonds in sample')

```

## Quiz6

Break out the histogram of diamond prices by cut ($cut).
You should have five histograms in separate
panels on your resulting plot.


```{r}
qplot(data = diamonds, x=price) +
  xlab('price') +
  ylab('occurrance')+
  facet_wrap(~cut)
```


## Quiz7

Which cut has the highest priced diamonds?
Which cut has the lowest priced diamonds?
Which cut has the lowest median diamonds?

```{r}
by(diamonds$price, diamonds$cut, summary)
```


This means the distributions should be somewhat similar,
but the histograms we created don't show that.

The 'Fair' and 'Good' diamonds appear to have 
different distributions compared to the better
cut diamonds. They seem somewhat uniform
on the left with long tails on the right.

Let's look in to this more.


## Quiz8


```{r}
qplot(data = diamonds, x=price) +
  xlab('price') +
  ylab('occurrance')+
  facet_wrap(~cut, scales="free")
```

# Quiz9
Create a histogram of price per carat
and facet it by cut. You can make adjustments
to the code from the previous exercise to get
started.

Adjust the bin width and transform the scale
of the x-axis using log10.

```{r}
qplot(data = diamonds, x=price, binwidth=0.05) +
  scale_x_log10() +
  xlab('log10 price') +
  ylab('occurrance')+
  facet_wrap(~cut, scales="free")
```
Did you find normal-ish distributions and a slightly bimodal distribution for Very Good cuts? (If the Very Good facet looks smooth, try decreasing your bin width! If the data looks skewed, did you remember to perform a log10 transform?)

#Quiz10

Investigate the price of diamonds using box plots,
numerical summaries, and one of the following categorical
variables: cut, clarity, or color.
```{r}
qplot(data = diamonds, y=price, geom="boxplot")
qplot(data = diamonds, y=price, x=cut, geom="boxplot")
qplot(data = diamonds, y=price, x=clarity, geom="boxplot")
qplot(data = diamonds, y=price, x=color, geom="boxplot")

```
```{r}
qplot(data = diamonds, x=cut)
qplot(data = diamonds, x=clarity)
qplot(data = diamonds, x=color)
```

#Quiz 11

What is the price range for the middle 50% of diamonds with color D AND J??
What is the IQR of diamonds with color D AND J??
```{r}
quantile(subset(diamonds, color=="D")$price)
quantile(subset(diamonds, color=="J")$price)

IQR(subset(diamonds, color=="D")$price)
IQR(subset(diamonds, color=="J")$price)
```

#Quiz12

Investigate the price per carat of diamonds across
the different colors of diamonds using boxplots.
```{r}
diamonds$price_per_carat <- diamonds$price / diamonds$carat

qplot(data = diamonds, y=price_per_carat, x=color, geom="boxplot")

```

# Quiz13

Investigate the weight of the diamonds in carat using a frquence polygon. use different bin widths to see how the frequency polygon changes. What carat size has a count greater than 2000?

```{r}
qplot(data = diamonds, x=carat, geom="freqpoly", binwidth=0.05)+
  scale_x_continuous(breaks=seq(0,5,0.2))


```
```{r}
table(diamonds$carat ==0.1)
table(diamonds$carat ==0.3)
table(diamonds$carat ==0.8)
table(diamonds$carat ==1.01)
table(diamonds$carat ==1.6)
table(diamonds$carat ==2.0)
table(diamonds$carat ==3.0)
table(diamonds$carat ==5.0)

```

#Quiz 15

dta manipulation with:

tidyr: A package that reshapes the layout of tables.
dplyr - a package that helps you transform tidy, tabular data
```{r}
#install.packages("tidyr", dependencies="T")
#library(tidyr)
#?gather
#?spread

#install.packages("dplyr", dependencies="T")
library(dplyr)
?select
?filter
?arrange
?mutate
?summarise
?group_by


```

